Exploring approaches ro probablistic record linkage using similarity searching

### 2017-10-30 paper.tex
Update on Overleaf.

\title{Efficient and accurate record linkage using similarity spaces}
%  \thanks{The authors would like to thank the Isaac Newton Institute
%  for Mathematical Sciences, Cambridge, for support and hospitality
%  during the programme \emph{Data Linkage and Anonymisation} where
%  this work was conducted (EPSRC grant EP/K032208/1). Peter Christen
%  was also supported by a grant from the Simons Foundation. The work
%  was also partially funded by the Australian Research Council
%  (DP130101801).}}

\author{Submitted for double-blind review}
%\author{Peter Christen\inst{1} \and Rainer Schnell\inst{2}
%        \and Dinusha Vatsalan\inst{1} \and Thilina
%        Ranbaduge\inst{1}}
%\institute{Research School of Computer Science,
%           The Australian National University, \\
%           Canberra, Australia.~
%           Contact: \email{peter.christen@anu.edu.au}
%           \and
%           Methodology Research Group, %Institute for Sociology,
%           University Duisburg-Essen, Duisburg, Germany.} %~
%%           \email{rainer.schnell@uni-due.de}}

@@

\keywords Entity resolution; deduplication; metric spaces; etc.

@@

We tried to develop a plan with PeterC. Here is the photo of the white-board.

\includegraphics[width=\textwidth]{2017-08-01_12_08_43.jpg}

@@

But, even better, here it is in a bulleted list.

@@

@@

@@

\begin{figure}[t]
  \centering
%  \includegraphics[width=1.0\textwidth]{xyz.eps}
  \caption{Descriptive caption
           %as outlined in Sect.~\ref{sec-overview} and detailed in
           %Sect.~\ref{sec-bf-attack}.
           \label{fig:xyz}}
\end{figure}

@@

\begin{figure}[t]
  \begin{center}
  \begin{scriptsize}
%    \begin{footnotesize}
  \begin{tabular}{ll} \hline
\multicolumn{2}{l}{\textbf{Algorithm 1: \emph{Candidate q-gram set
  generation}} -- Steps (1a) and (1b) } \\ \hline
\multicolumn{2}{l}{Input:} \\
\multicolumn{2}{l}{- $\mathbf{V}$: Set of attribute values and their
  frequencies from a public database} \\
\multicolumn{2}{l}{- $\mathbf{B}$: Set of BFs and their
  frequencies from the sensitive database} \\
\multicolumn{2}{l}{- $l$:~\, Length of Bloom filters} \\
\multicolumn{2}{l}{- $q$:~ Length of sub-strings to extract from
  attribute values} \\  
\multicolumn{2}{l}{- $m$: Minimum frequency for BFs and attribute
  values} \\
\noalign{\smallskip}
\multicolumn{2}{l}{Output:} \\
\multicolumn{2}{l}{- $\mathbf{C}$: List of possible q-grams for each
  BF position} \\
\noalign{\smallskip}
  1:  & $\mathbf{c}^+[p] = \{\}$, $\mathbf{c}^-[p] = \{\}$,
        $1 \le p \le l$ ~~~ // Initialize list of candidate q-gram
        sets \\
  2:  & $\mathbf{V}_F = \{v \in \mathbf{V}: v.f \ge m \}$
        ~~~~~~~~~~~~~~ // Attribute values with frequency of at
        least $m$ \\
  3:  & $\mathbf{B}_F = \{\mathbf{b} \in \mathbf{B}:
        \mathbf{b}.f \ge m \}$
        ~~~~~~~~~~~~~~\,// BFs with frequency of at least $m$ \\
  4:  & $revSort(\mathbf{B}_F)$, $revSort(\mathbf{V}_F)$ ~~~~~~~~~~ 
        // Sort according to frequencies, highest first \\
  5:  & $\mathbf{A} = [(\mathbf{b}_i, v_i) : 
        \mathbf{b}_i \in \mathbf{B}_F, v_i \in \mathbf{V}_F:
        \mathbf{b}_i.f > \mathbf{b}_j.f \land v_i.f > v_j.f:
        1 \le i < j \le min(|\mathbf{V}_F |, |\mathbf{B}_F|)]$ \\
  ~  &   ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ // Align $\mathbf{V}_F$
        and $\mathbf{B}_F$ as long as their frequencies are
        unique \\  
  6:  & \textbf{for} $(\mathbf{b}_i, v_i) \in \mathbf{A}$
        \textbf{do}: ~~~~~~~~~~~~~~~~~~~~ //  Step (1a): Get
        candidate sets of q-grams \\
  7:  & ~~~ $\mathbf{q}_i = genQGramSet(v_i, q)$ ~~~~~~~~\,//
        Convert attribute value into its q-gram set \\
  8:  & ~~~  \textbf{for} $1 \le p \le l$ \textbf{do}: 
        ~~~~~~~~~~~~~~~~~~~\, // Loop over all BF positions \\
  9:  & ~~~ ~~~ \textbf{if} $\mathbf{b}_i[p] == 1$ \textbf{then}:
        ~~~~~~~~~~~~\, // Bit at position p is 1 \\
  10: & ~~~ ~~~ ~~~ $\mathbf{c}^+[p] = \mathbf{c}^+[p] \cup
        \mathbf{q}_i$ ~~~~~~~~~~~\,// Add to set of possible
        q-grams \\
  11: & ~~~ ~~~ \textbf{else}: ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ //
        Bit at position $p$ is $0$ \\
  12: & ~~~ ~~~ ~~~ $\mathbf{c}^-[p] = \mathbf{c}^-[p] \cup
        \mathbf{q}_i$  ~~~~~~~~~~~// Add to set of \emph{not}
        possible q-grams \\
  13: & $\mathbf{C} = [\,]$ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,
        // Step (1b): Initialize empty list of q-gram sets \\
  14: & \textbf{for} $1 \le p \le l$ \textbf{do}: 
        ~~~~~~~~~~~~~~~~~~~~~~~~\,// Loop over all BF positions to
        combine q-gram sets \\
  15: & ~~~ $\mathbf{C}.append(\mathbf{c}^+[p] \setminus
        \mathbf{c}^-[p])$ ~~~~~~~~~~~\,// Remove not possible from
        possible q-grams \\
%  15: & ~~~ $\mathbf{c}[p] = \mathbf{c}^+[p] \setminus
%        \mathbf{c}^-[p]$ ~~~~~~~~~~~~~~~~~ // Remove not possible
%        from possible q-grams \\
%  16: & ~~~ $\mathbf{C}.append(\mathbf{c}[p])$
%        ~~~~~~~~~~~~~~~~~~~~~~ // Add to list of candidates for all
%        BF positions\\
16: & \textbf{return} $\mathbf{C}$ \\
  \hline
  \end{tabular}
%  \end{footnotesize}
 \end{scriptsize}
  \end{center}
  \label{algo:step1}
\end{figure}

@@

plot coverage as defined above for different data sets and the different blocking techniques

x-axis: similarity threshold from 0 to 1, y-axis is coverage as a
percentage (percentage of how many record pairs with a certain similarity value are generated by a blocking technique compared to all record pairs with that similarity - how to do this? we can't do full pair-wise comparison? can we get all pairs from m-tree for a certain distance, then run them through blocking?

For M-tree etc coverage will always be 1.0, while for other blocking techniques the coverage will be < 1.0 especially as similarities
become lower.

Plots should include average coverages for a certain similarity threshold over various parameter settings - e.g. for LHS or sorted neighbourhood, as well as min/max similarities (or std-dev bands in a transparent colour), as in the below example:

%http://www.randalolson.com/2014/06/28/how-to-make-beautiful-data-visualizations-in-python-with-matplotlib/

%% old text below  ***************************************

We conducted our experimental study using real data sets from two
domains.
%
The first are a pair of North Carolina Voter Registration (NCVR) data
sets (\url{ftp://alt.ncsbe.gov/data/})
%~\footnote{Available from \url{ftp://alt.ncsbe.gov/data/}}
collected in June 2014 (the database to be attacked) and October 2016
(the public database).
%, that have increasingly large differences in their time collection
%to see how this influences re-identification accuracy.
These data sets contain over five million records of voters including
their first names, surnames, and addresses. We present results of our
attack individually on the first name
%, last name, and address attributes, 
attribute, as well as on the concatenation of the first name and
surname attributes.

\begin{figure}[!t]
  \centering
%  \includegraphics[width=0.31\textwidth]
%  {reident-bf_len-ncvoter_20140619_1}
  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-hash_type-ncvoter_20140619_1}
  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-q-ncvoter_20140619_1}
  \\ ~ \\
%\includegraphics[width=0.31\textwidth]
%  {reident-bf_len-ncvoter_20140619_1-3}
  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-hash_type-ncvoter_20140619_1-3}
  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-q-ncvoter_20140619_1-3}
\caption{Results for the NCVR data sets with first name (top)
    and combination of first and surname (bottom) for different BF
    lengths (left), different hashing methods (middle), and different
    values of $q$ (right). No BF hardening was applied.
   \label{fig:ncvr}}
\end{figure}

The parameter settings we use in our experiments are $q = [2,3,4]$
(length of sub-strings used in BF encoding), BF length
$l = [250, 500, 1000,$ $2000]$, and either the double or random
hashing method~\cite{Sch16} as described in
Sect.~\ref{sec-overview}. We calculate the number of hash functions
$k$ based on $l$ and $q$ such that the false positive rate is
minimized~\cite{Vat14c}. We also apply the BF hardening techniques
balancing and XOR-folding~\cite{Sch16}.
% as described in Sect.~\ref{sec-overview}.
We use different numbers for the most frequent attribute values
(ranked by their frequencies) to be re-identified: $|\mathbf{G}| =
[10,20,50,100]$.
%most frequent values from the public data set to be guessed.

\textbf{Discussion:}~
In Fig.~\ref{fig:ncvr} we show the results for the NCVR data sets.
% for one individual attribute (first name) in the top row, and
%three individual attributes (first name, last name, and address in
%the first three rows, respectively), and
%combined attributes of first and surnames in the bottom row.
% BFs are hardened using the balanced method.
%
As can be seen, when the BF length $l$ increases (left column) the
percentage of correct re-identifications mostly increases, as with
larger $l$ the number of q-grams mapped to a certain bit position
decreases. All values can be correctly re-identified for individual
attributes
%nearly $90\%$ of values can be correctly re-identified
%%as 1-to-1 or 1-to-many matches
when the number of frequent values is 10.
%or 20. 
Around half of all
values can still be re-identified even when values from two
attributes

\begin{table}[!t]
 %\addtolength{\tabcolsep}{-3pt}
  \caption{Comparison of re-identification results with existing BF
     attack methods.}
     \label{table:exist_approaches}
  \centering
  \begin{scriptsize}
    \begin{tabular}{llccc}
    \hline\noalign{\smallskip}
   Publication & Data set & Num BFs~ & ~1-1 corr~ & Run time
     \\ \noalign{\smallskip}\hline\noalign{\smallskip}
  Kuzu et al. (2011)~\cite{Kuz11} & NCVR first names & 3,500 & 400 &
    1,000 sec \\
  Kuzu et al. (2013)~\cite{Kuz13} & Patient names & 20 & 4 & 
    few sec \\
  ~~~~ " & ~~~~ " & 42 & 0 & $>$ week \\
  Niedermeyer et al. (2014)~\cite{Nie14}~ & German surnames & 7,580 &
    934 & $>$ days  \\
  Kroll and Steinmetzer (2015)~\cite{Kro15}~ & German names and
    locations & 100K & 44K & $>$ days \\
  Our approach & NCVR first names & 10--100 & 7--10 & 0.73--0.75 sec
    \\
  ~~~~ " & NCVR first and surnames & 10--100 & 3--6 & 1.5--1.9
  sec  \\
%  ~~~~ " & UKCD surname and address~ & 10 & 10 & 0.05 sec \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

@@

%\bibliographystyle{abbrv}
\bibliographystyle{splncs03}
\bibliography{paper}

>>>>

\title{Exploring approaches ro probablistic record linkage using similarity searching}
%  \thanks{The authors would like to thank the ...|

\author{Submitted for double-blind review}
%\author{Özgür Akgün\inst{1} and Alan Dearle\inst{1} and Graham Kirby\inst{1} Peter Christen\inst{2}

%\institute{School of Computer Science,
%           University of St Andrews, \\
%           St Andrews, Scotland.~
%           Contact: \email{ozgur.akgun@st-andrews.ac.uk}

%\institute{Research School of Computer Science,
%           The Australian National University, \\
%           Canberra, Australia.~
%           Contact: \email{peter.christen@anu.edu.au}

@@

\keywords Probablistic record Linkage; similarity search; metric spaces; etc.

@@

@@

Plan in a bulleted list:

@@

Note: The rest of this paper should be regarded as a placeholder for now.

@@

The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977. This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records. As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.

@@

\section{Datasets}
\label{sec-data}

We conducted our experimental study using four real data sets from three
domains. The first is the CORA data set  (\url{https://linqs-data.soe.ucsc.edu/public/lbc/cora.tgz}) \textbf{ which consists of 2708 scientific publications classified into one of seven classes. The citation network consists of 5429 links.}

%Fields and true links here

The second are a pair of sets of records drawn from the North Carolina Voter Registration (NCVR) data
sets (\url{ftp://alt.ncsbe.gov/data/}) collected in June 2014 and October 2016
 \textbf{( who collected? where is the dataset published?)}.
These data sets contain over five million records of voters encoded into 11 fields including
their first names, surnames, and addresses. Ground truth is provided via an \textit{voter\_id} field which is invariant across the data sets.
In our experiments we use two subsets of this data comprising 224074 and 224062 records drawn from the original five million plus records. \textbf{chosen how?}

We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets of the Digitising Scotland data set. One data set contains records of vital events registered on the Isle of Skye, a rural district, while the other contains records from Kilmarnock, an industrial town. Both cover the period 1861-1901. These are different types of communities, with different family structures and name distributions \cite{reid2002}. Both datasets are encoded with the same schemata and include names, sex, addresses and the names of each person's father and mother. Ground truth is provided using demographer encoded identifiers which provide (incomplete) linkage between the records.  The Isle of Skye dataset comprises around 17,600 birth records, 12,300 death records, and 2,700 marriage records.
The demographers  identified:
\begin{itemize}
\item 4,300 family groups of siblings, with a mean family size of 3.9 siblings and a maximum family size of 16,
\item 2,900 links from a birth record to the child's death record,
\item 60 links from a marriage record to the groom's death record, and
\item 100 links from a marriage record to the bride's death record.
\end{itemize}
The Kilmarnock data set contains around 38,400 birth records, 23,700 death records, and 8,700 marriage records. Here, the demographers identified:
\begin{itemize}
\item 13,100 family groups of siblings, with a mean family size of 2.8 siblings and a  maximum family size of 16,
\item 8,300 links from a birth record to the child's death record,
\item 30 links from a marriage record to the groom's death record, and
\item 50 links from a marriage record to the bride's death record.
\end{itemize}

@@

@@

@@

\bibliographystyle{abbrv}
\bibliography{paper.bib}
